---
title: "Study_try"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
```

# Meta-Analytic Data {.tabset}

## Install and load relevant `R` packages

```{r Packages, include = TRUE, results = 'hide', message=FALSE}
# Install R packages (if needed)
# install.packages("pacman")
library(pacman)
pacman::p_load(psych,
               metaSEM,
               metafor,
               robumeta,
               clubSandwich,
               ggplot2,
               POMADE,
               dplyr)

## Number of available cores
ncores <- parallel::detectCores() - 1

## Increase the output length for display
options(max.print = 10000)

```

## Data Input and Preparation
```{r Data input, include = TRUE, cache = TRUE}
## Data set in wide format
ageyouth <- read_xlsx("output/processed_data_aw.xlsx")
## Check the data
head(ageyouth, 10)

# --- Basic preparation ---
ageyouth <- ageyouth %>%
  rename(
    StudyID   = `1_ID`,
    SampleID  = `1.1_CaseID`,
    ESID      = RowID,
    n         = `10_Size`,
    age       = `11_Mage`,
    female    = `18_Female`,
    construct = `Negative/Positive+LS`,
    income    = income_level,
    m_i       = WBmi,
    sd_i      = WBsdi,
    m_j       = WBmj,
    sd_j      = WBsdj,
    r_ij      = WBcorrij,
    timelag   = timelagc
  ) %>%
  mutate(
    n = as.numeric(n),
    age = as.numeric(age),
    female = ifelse(female > 1, female/100, female),
    r_ij = as.numeric(r_ij),
    timelag = as.numeric(timelag),
    construct = as.factor(construct),
    income = as.factor(income)
  )
```
## Description of the Meta-Analytic Dataset

```{r data description, include = TRUE, message=FALSE, cache = TRUE}
## Number of effect sizes
nrow(ageyouth)

## Number of articles
length(table(ageyouth$StudyID))
table(ageyouth$StudyID)

## Average number of effect sizes per study
mean(table(ageyouth$StudyID)/10)
sd(table(ageyouth$StudyID)/10)
median(table(ageyouth$StudyID)/10)
min(table(ageyouth$StudyID)/10)
max(table(ageyouth$StudyID)/10)

## Number of samples
length(table(ageyouth$SampleID))
table(ageyouth$SampleID)

## Overall sample size
OverallN <- round(round(sum(aggregate(ageyouth$n, 
                                      by = list(ageyouth$SampleID),
                                      FUN = mean)$x, 
                            na.rm = TRUE), 
                        0))
OverallN

## Distribution of sample sizes
psych::describe(ageyouth$n)
hist(ageyouth$n,
     xlab = "Sample size",
     main = "Distribution of the study sample sizes")

## Effect sizes
psych::describe(ageyouth$es_dyear)
hist(ageyouth$es_dyear,
     xlab = "Standardized mean difference (corrected by year)",
     main = "Distribution of the effect sizes")

```

# Compute effect sizes

```{r A. Mean-level change (SMCR)}

es_ml <- escalc(
  measure = "SMCR",
  m1i = m_j, sd1i = sd_j,
  m2i = m_i, sd2i = sd_i,
  ni  = n, ri = r_ij,
  data = dat
) %>%
  mutate(type = "mean_level")

## B. Sequential correlations (Fisher z)
es_sc <- escalc(
  measure = "ZCOR",
  ri = r_ij, ni = n,
  data = dat
) %>%
  mutate(type = "seq_corr")

# Combine both
es <- bind_rows(es_ml, es_sc) %>%
  mutate(
    age_c = scale(age, center = TRUE, scale = FALSE) %>% as.numeric(),
    age2 = age_c^2,
    age3 = age_c^3
  )
```

## Helper function for CHE-style model (From Ronny's Code)

```{r}

fit_CHE <- function(df, rho = 0.6) {
  V <- vcalc(
    vi = df$vi,
    cluster = df$StudyID,
    subgroup = df$SampleID,
    obs = df$ESID,
    rho = rho,
    data = df
  )
  model <- rma.mv(
    yi = yi, V = V, data = df,
    random = ~ 1 | StudyID/ESID,
    method = "REML",
    sparse = TRUE
  )
  list(model=model, V=V)
}


### Same ID from Ronny's code

# Fit base models 
ml_data <- es %>% filter(type == "mean_level")
sc_data <- es %>% filter(type == "seq_corr")

m_ml <- fit_CHE(ml_data)
m_sc <- fit_CHE(sc_data)

cat("\n Base Model: Mean-level (SMCR) \n")
print(summary(m_ml$model))
print(coef_test(m_ml$model, vcov="CR2", cluster=ml_data$StudyID))

cat("\n Base Model: Sequential (Fisher z) \n")
print(summary(m_sc$model))
print(coef_test(m_sc$model, vcov="CR2", cluster=sc_data$StudyID))
```
# Sample Study with Moderator Age, Female, Income, Timelag

```{r}
#  Moderator models 
fit_mod <- function(df, rho = 0.6) {
  V <- vcalc(
    vi = df$vi,
    cluster = df$StudyID,
    subgroup = df$SampleID,
    obs = df$ESID,
    rho = rho,
    data = df
  )
  fm <- rma.mv(
    yi ~ construct + age_c + age2 + age3 + female * income + timelag,
    V = V, data = df,
    random = ~ 1 | StudyID/ESID,
    method = "REML",
    sparse = TRUE
  )
  list(
    model = fm,
    robust = coef_test(fm, vcov="CR2", cluster=df$StudyID)
  )
}

mod_ml <- fit_mod(ml_data)
mod_sc <- fit_mod(sc_data)

cat("\n=== Moderator Model: Mean-level ===\n")
print(summary(mod_ml$model))
print(mod_ml$robust)

cat("\n=== Moderator Model: Sequential ===\n")
print(summary(mod_sc$model))
print(mod_sc$robust)
```
## 1. Moderator analyses For Construct (Positive/Negative Affect)

```{r}
U <- metafor::vcalc(vi = v_dyear,            # Sampling variances
                    cluster = StudyID,       # Hierarchical level: Studies
                    subgroup = SampleID,     # Correlational level: Samples
                    obs = ESID,              # Level of effect sizes
                    type = factor(Negative.Positive.LS+1),  # Different constructs
                    data = agewell,          # Data set
                    rho = list(rho, rho),      # Within-sample correlations of effects per construct
                    checkpd = TRUE, 
                    nearpd = FALSE)  
```


```{r che model study moderator, include = TRUE, cache = TRUE}

## Positive vs. negative LS
table(ageyouth$Negative.Positive.LS)

## Model specification
## Direct moderator effects
CHE.model.mod <- rma.mv(yi = es_dyear,
                    V = v_d,
                    data = ageyouth,
                    random = list(~ 1 | StudyID/ESID),
                    method = "REML",
                    time = TRUE,
                    sparse = TRUE,
                    mods =~ construct)

## Model summary
summary(CHE.model.mod)

## Cluster-robust standard errors and p-values
## with small-sample adjustment
conf_int(CHE.model.mod, vcov = "CR2")
coef_test(CHE.model.mod, vcov = "CR2")

## Variance explanation (R-squared)
## Note: This is the proportional reduction of variances.
## Study level
max(0, 1- CHE.model.mod$sigma2[1]/CHE.model$sigma2[1])
## Effect size level (within studies)
max(0, 1- CHE.model.mod$sigma2[2]/CHE.model$sigma2[2])




## Model specification
## Effects per moderator category
## See also: https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/multilevel-ma.html#three-level-subgroup
CHE.model.mod2 <- rma.mv(yi = es_dyear,
                    V = v_d,
                    data = ageyouth,
                    random = list(~ 1 | StudyID/ESID),
                    method = "REML",
                    time = TRUE,
                    sparse = TRUE,
                    mods =~ factor(construct) - 1) # turn off the intercept

## Model summary
summary(CHE.model.mod2)

```
# above all everyhing should be similar

# 2. Moderator Analysis: Age: linear, quadratic, cubic 

```{r}
## AGE (linear, quadratic, cubic)
## Center age before adding powers
ageyouth$agem1ic1 <- scale(ageyouth$age, center = TRUE, scale = FALSE)
ageyouth$agem1ic2  <- ageyouth$agem1ic1^2
ageyouth$agem1ic3  <- ageyouth$agem1ic1^3

## Model specification
CHE.model.age <- rma.mv(yi = es_dyear,
                        V = v_d,
                        data = ageyouth,
                        random = list(~ 1 | StudyID/ESID),
                        method = "REML",
                        time = TRUE,
                        sparse = TRUE,
                        mods = ~ agem1ic1 + agem1ic2 + agem1ic3)

## Model summary
summary(CHE.model.age)

## Cluster-robust SEs and p-values
conf_int(CHE.model.age, vcov = "CR2")
coef_test(CHE.model.age, vcov = "CR2")

## Variance explanation (R?)
max(0, 1 - CHE.model.age$sigma2[1]/CHE.model$sigma2[1])   # study level
max(0, 1 - CHE.model.age$sigma2[2]/CHE.model$sigma2[2])   # effect-size level

```

# 3. Moderator Analysis: Female + Country Income

```{r}
## FEMALE + COUNTRY INCOME
## (Main effects + their interaction if desired)

CHE.model.sex.inc <- rma.mv(yi = es_dyear,
                            V = v_d,
                            data = ageyouth,
                            random = list(~ 1 | StudyID/ESID),
                            method = "REML",
                            time = TRUE,
                            sparse = TRUE,
                            mods = ~ female + factor(income))

## Model summary
summary(CHE.model.sex.inc)

## Cluster-robust SEs and p-values
conf_int(CHE.model.sex.inc, vcov = "CR2")
coef_test(CHE.model.sex.inc, vcov = "CR2")

## Variance explanation (R?)
max(0, 1 - CHE.model.sex.inc$sigma2[1]/CHE.model$sigma2[1])
max(0, 1 - CHE.model.sex.inc$sigma2[2]/CHE.model$sigma2[2])

## --- Optional: add interaction term ---
CHE.model.sex.inc.int <- rma.mv(yi = es_dyear,
                                V = v_d,
                                data = ageyouth,
                                random = list(~ 1 | StudyID/ESID),
                                method = "REML",
                                time = TRUE,
                                sparse = TRUE,
                                mods = ~ female * factor(income))

summary(CHE.model.sex.inc.int)
conf_int(CHE.model.sex.inc.int, vcov = "CR2")
coef_test(CHE.model.sex.inc.int, vcov = "CR2")

```

# 4. Moderator Analysis: Time Lag

```{r}
## TIME LAG
## (assumes column timelagc already centered or numeric)

CHE.model.lag <- rma.mv(yi = es_dyear,
                        V = v_d,
                        data = ageyouth,
                        random = list(~ 1 | StudyID/ESID),
                        method = "REML",
                        time = TRUE,
                        sparse = TRUE,
                        mods = ~ timelag)

## Model summary
summary(CHE.model.lag)

## Cluster-robust SEs and p-values
conf_int(CHE.model.lag, vcov = "CR2")
coef_test(CHE.model.lag, vcov = "CR2")

## Variance explanation (R?)
max(0, 1 - CHE.model.lag$sigma2[1]/CHE.model$sigma2[1])
max(0, 1 - CHE.model.lag$sigma2[2]/CHE.model$sigma2[2])

```

```{r session}
sessionInfo()
```






