---
title: "Supplementary Material"
author: "Meta-Analytic Baseline Model with Correlated and Hierarchical Effects"
date: "23 September 2025"
fontsize: 12pt
geometry: margin = 1in
urlcolor: blue
output: 
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float: TRUE
    number_sections: FALSE
    theme: default
    highlight: default
    

---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Meta-Analytic Data {.tabset}

## Install and load relevant `R` packages

```{r Packages, include = TRUE, results = 'hide', message=FALSE}
# Install R packages (if needed)
# install.packages("pacman")
library(pacman)
pacman::p_load(psych,
               metaSEM,
               metafor,
               robumeta,
               clubSandwich,
               ggplot2,
               POMADE,
               dplyr)

## Number of available cores
ncores <- parallel::detectCores() - 1

## Increase the output length for display
options(max.print = 10000)

```


## Data Input and Preparation
```{r Data input, include = TRUE, cache = TRUE}
## Data set in wide format
agewell <- read.csv2("data_complete_rs.csv", header = TRUE)
## Check the data
head(agewell, 10)


```


## Description of the Meta-Analytic Dataset

```{r data description, include = TRUE, message=FALSE, cache = TRUE}
## Number of effect sizes
nrow(agewell)

## Number of articles
length(table(agewell$StudyID))
table(agewell$StudyID)

## Average number of effect sizes per study
mean(table(agewell$StudyID)/10)
sd(table(agewell$StudyID)/10)
median(table(agewell$StudyID)/10)
min(table(agewell$StudyID)/10)
max(table(agewell$StudyID)/10)

## Number of samples
length(table(agewell$SampleID))
table(agewell$SampleID)

## Overall sample size
OverallN <- round(round(sum(aggregate(agewell$X10_Size, 
                                      by = list(agewell$SampleID),
                                      FUN = mean)$x, 
                            na.rm = TRUE), 
                        0))
OverallN

## Distribution of sample sizes
psych::describe(agewell$X10_Size)
hist(agewell$X10_Size,
     xlab = "Sample size",
     main = "Distribution of the study sample sizes")

## Effect sizes
psych::describe(agewell$es_dyear)
hist(agewell$es_dyear,
     xlab = "Standardized mean difference (corrected by year)",
     main = "Distribution of the effect sizes")

```


# Constructing Sampling Covariance Matrices ($V_i$)

## Overall data
As a first step, we construct the sampling variance-covariance matrices. These matrices contain the known sampling variances in the diagonal and the sampling covariances in the off-diagonal part.

```{r sampling cov matrices, include = TRUE, cache = TRUE}

## Generate the sampling covariance matrices
## Note: Each sample can contribute with multiple effect sizes.
## Hence, SampleID identifies the correlational dependencies (e.g., multiple
## measures or time points).

## Set a within-sample correlation
rho <- 0.6

## Source: https://wviechtb.github.io/metafor/reference/vcalc.html
V <- metafor::vcalc(vi = v_dyear,            # Sampling variances
                    cluster = StudyID,       # Hierarchical level: Studies
                    subgroup = SampleID,     # Correlational level: Samples
                    obs = ESID,              # Level of effect sizes
                    data = agewell,          # Data set
                    rho = rho,               # Within-sample correlation of effects
                    checkpd = TRUE,          # Check for non-positive definiteness
                    nearpd = TRUE)           # Replace non-pd matrices by nearest pd matrices

## Illustrate some matrices V for some samples
blsplit(V, agewell$StudyID)$`11588` # One sample measured over six time points
blsplit(V, agewell$StudyID)$`11591` # Two samples measured over three time points

## Check for non-positive definite covariance matrices
metaSEM::is.pd(blsplit(V, agewell$StudyID), cor.analysis = FALSE)
metaSEM::is.pd(blsplit(V, agewell$SampleID), cor.analysis = FALSE)
## Result: All covariance matrices for samples and studies are positive definite (good!).

```


## Alternative: Construct-specific data
While the former code did not distinguish between different constructs or types of outcomes, the following code does this.

```{r sampling cov matrices2, include = TRUE, cache = TRUE}

## Source: https://wviechtb.github.io/metafor/reference/vcalc.html
U <- metafor::vcalc(vi = v_dyear,            # Sampling variances
                    cluster = StudyID,       # Hierarchical level: Studies
                    subgroup = SampleID,     # Correlational level: Samples
                    obs = ESID,              # Level of effect sizes
                    type = factor(Negative.Positive.LS+1),  # Different constructs
                    data = agewell,          # Data set
                    rho = list(rho, rho),      # Within-sample correlations of effects per construct
                    checkpd = TRUE, 
                    nearpd = FALSE)    

## Illustrate some matrices V for some samples
blsplit(U, agewell$StudyID)$`11588` # One sample measured over six time points
blsplit(U, agewell$StudyID)$`11591` # Two samples measured over three time points

```



# Correlated and Hierarchical Effect Size (CHE) Models

## Model with study level and without moderators

```{r che model study, include = TRUE, cache = TRUE}

## Reference: Pustejovsky & Tipton (2022), https://doi.org/10.1007/s11121-021-01246-3
## Source: https://wviechtb.github.io/metafor/reference/rma.mv.html

## Model specification
CHE.model <- rma.mv(yi = es_dyear,
                    V = V,
                    data = agewell,
                    random = list(~ 1 | StudyID/ESID),
                    method = "REML",
                    time = TRUE,
                    sparse = TRUE)

## Model summary
summary(CHE.model)

## Confidence intervals of variances
## Note: This can take some time.
#confint.rma.mv(CHE.model)

## Cluster-robust standard errors
CHE.model.robust <- robust(CHE.model, 
                           cluster = StudyID, 
                           clubSandwich = TRUE)
summary(CHE.model.robust)

## Processing time in seconds
CHE.model$time

## Model fit
metafor::fitstats(CHE.model)

## Alternative specification (for model comparisons):
aCHE.model <- rma.mv(yi = es_dyear,
                    V = V,
                    data = agewell,
                    random = list(~ 1 | ESID,
                                  ~ 1 | StudyID),
                    method = "REML",
                    time = TRUE,
                    sparse = TRUE)

summary(aCHE.model)

## I-squared values
## Source: https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate
W <- diag(1/CHE.model$vi)
X <- model.matrix(CHE.model)
P <- W - W %*% X %*% solve(t(X) %*% W %*% X) %*% t(X) %*% W
## Overall I-squared
100 * sum(CHE.model$sigma2) / (sum(CHE.model$sigma2) + (CHE.model$k-CHE.model$p)/sum(diag(P)))
## Level-specific I-squared
100 * CHE.model$sigma2 / (sum(CHE.model$sigma2) + (CHE.model$k-CHE.model$p)/sum(diag(P)))

```


## Model with study and sample level and without moderators

```{r che model sample study, include = TRUE, cache = TRUE}

## Reference: Pustejovsky & Tipton (2022), https://doi.org/10.1007/s11121-021-01246-3
## Source: https://wviechtb.github.io/metafor/reference/rma.mv.html

## Model specification
CHE.model2 <- rma.mv(yi = es_dyear,
                    V = V,
                    data = agewell,
                    random = list(~ 1 | ESID,
                                  ~ 1 | SampleID,
                                  ~ 1 | StudyID),
                    method = "REML",
                    time = TRUE,
                    sparse = TRUE)

## Model summary
summary(CHE.model2)

## Confidence intervals of variances
## Note: This can take some time.
#confint.rma.mv(CHE.model2)

## Cluster-robust standard errors
CHE.model2.robust <- robust(CHE.model2, 
                           cluster = StudyID, 
                           clubSandwich = TRUE)
summary(CHE.model2.robust)

## Processing time in seconds
CHE.model2$time

## Model fit
metafor::fitstats(CHE.model2)

## Model comparison: aCHE.model vs. CHE.model2
metafor::anova.rma(aCHE.model, CHE.model2)
## Result: Model aCHE.model fits better (lower information criteria), and
## the heterogeneity at the level of samples is zero.

## I-squared values
## Source: https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate
W2 <- diag(1/CHE.model2$vi)
X2 <- model.matrix(CHE.model2)
P2 <- W2 - W2 %*% X2 %*% solve(t(X2) %*% W2 %*% X2) %*% t(X2) %*% W2
## Overall I-squared
100 * sum(CHE.model2$sigma2) / (sum(CHE.model2$sigma2) + (CHE.model2$k-CHE.model2$p)/sum(diag(P2)))
## Level-specific I-squared
100 * CHE.model2$sigma2 / (sum(CHE.model2$sigma2) + (CHE.model2$k-CHE.model2$p)/sum(diag(P2)))

```


## Model with sample level and without moderators

```{r che model sample, include = TRUE, cache = TRUE}

## Reference: Pustejovsky & Tipton (2022), https://doi.org/10.1007/s11121-021-01246-3
## Source: https://wviechtb.github.io/metafor/reference/rma.mv.html

## Model specification
CHE.model3 <- rma.mv(yi = es_dyear,
                    V = V,
                    data = agewell,
                    random = list(~ 1 | SampleID/ESID),
                    method = "REML",
                    time = TRUE,
                    sparse = TRUE)

## Model summary
summary(CHE.model3)

## Confidence intervals of variances
## Note: This can take some time.
#confint.rma.mv(CHE.model3)

## Cluster-robust standard errors
CHE.model3.robust <- robust(CHE.model3, 
                           cluster = StudyID, 
                           clubSandwich = TRUE)
summary(CHE.model3.robust)

## Processing time in seconds
CHE.model3$time

## Model fit
metafor::fitstats(CHE.model3)

## I-squared values
## Source: https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate
W3 <- diag(1/CHE.model3$vi)
X3 <- model.matrix(CHE.model3)
P3 <- W3 - W3 %*% X3 %*% solve(t(X3) %*% W3 %*% X3) %*% t(X3) %*% W3
## Overall I-squared
100 * sum(CHE.model3$sigma2) / (sum(CHE.model3$sigma2) + (CHE.model3$k-CHE.model3$p)/sum(diag(P3)))
## Level-specific I-squared
100 * CHE.model3$sigma2 / (sum(CHE.model3$sigma2) + (CHE.model3$k-CHE.model3$p)/sum(diag(P3)))

```


## Decision on a working model

Given that heterogeneity mainly exists within and between studies, yet not at the level of samples, we decide for the model `CHE.model` as the meta-analytic working model.


## Power analysis of the working model (`CHE.model`)

The R package `POMADE` allows for the analysis of statistical power with such complex models.

```{r che model power, include = TRUE, cache = TRUE}

## Reference: Vembye et al. (2024), https://doi.org/10.1002/jrsm.1752
## Source: https://mikkelvembye.github.io/POMADE/

## Calculating power
power_MADE(
  J = length(table(agewell$StudyID)),   # Number of primary studies
  mu = CHE.model$beta[,1],    # Effect size
  tau = CHE.model$sigma2[1],            # Between-study variation
  omega = CHE.model$sigma2[2],          # Within-study variation
  rho = rho,                        # Correlations among effect sizes
  model = "CHE",                    # Type of working model
  var_df = "RVE",                   # Variance estimation
  sigma2_dist = aggregate(agewell$v_dyear, 
                          by = list(agewell$StudyID), 
                          FUN = mean)$x,  # Average sampling variance per study
  n_ES_dist = as.data.frame(table(agewell$StudyID))$Freq,   # Number of effect sizes per study
  iterations = 1000,
  seed = 3617
)

## Result: We have 100% power to detect the effect we have actually found.


```


## Moderator analyses

```{r che model study moderator, include = TRUE, cache = TRUE}

## Positive vs. negative LS
table(agewell$Negative.Positive.LS)

## Model specification
## Direct moderator effects
CHE.model.mod <- rma.mv(yi = es_dyear,
                    V = V,
                    data = agewell,
                    random = list(~ 1 | StudyID/ESID),
                    method = "REML",
                    time = TRUE,
                    sparse = TRUE,
                    mods =~ Negative.Positive.LS)

## Model summary
summary(CHE.model.mod)

## Cluster-robust standard errors and p-values
## with small-sample adjustment
conf_int(CHE.model.mod, vcov = "CR2")
coef_test(CHE.model.mod, vcov = "CR2")

## Variance explanation (R-squared)
## Note: This is the proportional reduction of variances.
## Study level
max(0, 1- CHE.model.mod$sigma2[1]/CHE.model$sigma2[1])
## Effect size level (within studies)
max(0, 1- CHE.model.mod$sigma2[2]/CHE.model$sigma2[2])


## Model specification
## Effects per moderator category
## See also: https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/multilevel-ma.html#three-level-subgroup
CHE.model.mod2 <- rma.mv(yi = es_dyear,
                    V = V,
                    data = agewell,
                    random = list(~ 1 | StudyID/ESID),
                    method = "REML",
                    time = TRUE,
                    sparse = TRUE,
                    mods =~ factor(Negative.Positive.LS) - 1) # turn off the intercept

## Model summary
summary(CHE.model.mod2)

```



# Further analyses

Given that the within-sample correlation (due to multiple measures and/or time points) is set to an arbitrary value, sensitivity analyses varying this correlation ($\rho$) are needed.

Some analyses of publication bias should also be based on this working model (e.g., PET-PEESE).

 
# R session info
```{r, include = TRUE, cache = TRUE}
sessionInfo()
```